Numerical simulation of bubble deformation \\ and breakup under simple linear shear flows

### 2024-04-01 Reply-Reviewer2.tex

%%7
\par\noindent
\item
\textsf
Page 6: It is well-known that interface capturing methods have drawbacks in
terms of accuracy, which in this problem could lead to uncertainty or
inaccuracy concerning the breakup conditions and the related bubble shapes.
Indeed, the topological changes may affect the small characteristic size of the
filament formed by the distorted drop or bubble, when it becomes comparable
with the grid size. For instance, in their study of two bubbles rising in line
(using VOF and Basilisk as a solver), Zang and Magnaudet, ref. [5], found that
refinement up to R/272 (i.e. the local grid size is 272 times smaller than the
bubble radius) is required in the proximity of the interface to properly
capture the topological changes of millimetric bubbles, since only then the
characteristic size of the grid is smaller than the average film thickness in
typical coalescence conditions. Instead, the authors use R/24 local grid size
at the most. Moreover, the Adaptive Mesh Refinement (AMR) technique is usually
employed for 3-4 consecutive levels of refinement, at least. In the present
work, only two levels of refinement have been used, why? The mesh convergence
study seems quite incomplete.  A more complete study with figures to verify it
should be included in an appendix concerning the mesh and time step
independence of the results. The type of time discretization used is not
stated; is it explicit and which algorithm is used?
\vspace{3 mm}

Response: \\
\par\noindent
In response to the question ``In the present
work, only two levels of refinement have been used, why?'', we added the 
following text to our revision, (under the heading ``selecting the 
appropriate grid size'')
\par\noindent

{ \color{red}
The grid size and adaptive meshing strategy that we adopt is chosen
in order to answer the research question as to the conditions which determine
whether a bubble in shear flow
will break-up or not.  In such a case, we must accurately capture the 
balance of forces with respect to the (non-local) force exerted from the 
wall driven 
flow acting against the interfacial
surface tension force.   
The accuracy of the ``Critical Reynolds Number''
depends on the 
largest Taylor Deformation parameter $D$
that is supported by the grid (see e.g. Figures 8 and 11).  
As we report here, we have found that as
long as the grid size is fine enough to support a Taylor Deformation parameter
$D<0.95$, then the transition region 
(i.e. ``Critical Reynolds number'')
(see Figures 3 and 12) will be captured
with a tolerance of three percent.  The simulation time becomes impractical
if we were to try to further improve the ``critical Reynolds number''
accuracy.  
A smaller tolerance
would necessitate a larger supported Deformation
parameter $D$ which would in turn necessitate a higher aspect ratio
computational domain, increased droplet surface area at break-up,
increased number of time steps, and higher resolution for representing the
drop/bubble at its thinnest point.

We make the distinction between our present research, and the research found
in the work of Zang, Ni, 
and Magnaudet[47,48] on predicting the
conditions for bubble mergers.  Even in the most extreme cases for
mergers, the largest Deformation parameter never exceeds
$0.4$ in [47].  In summary, our gridding requirements 
necessitate grid points distributed relatively evenly throughout the 
computational domain when a bubble is stretched to a $0.9$ Deformation,
whereas in [47] the gridding strategy necessitates 
a more localized strategy.

@@

%%8
\par\noindent
\item
\textsf
{\color{red} Several experimental studies concerning bubbles in shear flow
exist (for example, ref. [6]) so alternatively one could compare the numerical
results with experiments and avoid the comparison with other numerical results,
which are older and may not have achieved convergence due to lack of access to
the software and hardware we have today.}
\vspace{3 mm}

Response: \\
{\color{red} 
%As a quantitative comparison, we have newly made a comparison for bubble 
%deformation as Table I\hspace{-1.2pt}I:
We added a new quantitative comparison for bubble deformation in Table
I\hspace{-1.2pt}I to verify the validity of our computational results; in the
table we compare against the experimental results of Muller-Fischer et al.
(Experiments in fluids, 2008) and Rust and Manga (Journal of Colloid and
Interface Science, 2002). The new numerical results for bubble deformation
indicate that our computations are consistent with past studies.
%We believe that the effectivity of our computations could be proved for bubble deformation.
}
\\

%%9
\par\noindent
\item
\textsf
{The following is the most disturbing issue with this study: It is fine to
examine conditions leading to bubble breakup or other flow instabilities, but
the range of parameter values investigated must correspond to existing
materials and achievable flow conditions.  The authors have given the governing
equations in dimensional form and defined dimensionless numbers incorrectly as
stated in point \#2, but also assigned them values that may not be physically
relevant. Moreover, the bubble size is nowhere stated.  \\
Finally, the Reynolds number of the channel in this case would be: $Re_{\rm
channel} \simeq 6000$.  This high value of Re corresponds to turbulence
conditions.}
\vspace{3 mm}

Response: \\
As we have stated in response to your comment \#2, correct dimensionless
numbers based on the density and the viscosity of the surrounding liquid were
used in real computations.  Regarding physical properties and conditions in our
computations, we will give you an example for the case of $Ca$ = 0.3 and $Re$
=93.\\
\par\noindent
Physical properties 
(a silicone oil has a very similar physical properties with these values):\\
\quad $\rho_{\rm m}$ (density of liquid) = 1000 kg/m$^3$, 
$\mu_{\rm m}$ (viscosity of liquid) = 0.2 Pa$\cdot$s,
$\sigma$ (surface tension) = 0.025 N/m, $R$ (bubble radius) = 5.0 mm.\\
From ${\it \Gamma}$ = $2V/H$ = $2V/6R$, \  
$V$ = 3${\it \Gamma} R$ = $\frac{3\sigma Ca}{\mu_{\rm m}} $ = 1.1 m/s.
Also, we obtain $Fr$ (Froude number) = 
$\frac{{\it \Gamma}R}{\sqrt {gR}}$ = 1.7. \\
In this flow system, the effect of gravity may not be completely 
negligible. However, in our computations, 
we set $g = 0$ because we wanted to clearly isolate only the effects of 
the density and viscosity ratios. 
From another perspective, it can be said that we investigated 
phenomena in space.
At the end of the Problem Description section, 
we added a detailed explanation about physical properties and conditions.

@@

Also, regarding the channel $Re$ number, we obtain $Re_{\rm channel} =
\frac{\rho_{\rm m}2VH}{\mu_{\rm m}}$ = 3300 for a simple linear shear-flow.  As
expected, $Re_{\rm channel} = 3300$ falls within turbulence conditions for
simple linear shear-flow of a single phase fluid.  It is important to note that
the generation of turbulence is suppressed and turbulent eddy dissipation
becomes large if bubbles exist in fluid flows. However, as far as we know,
there are no conclusive results on turbulence in a simple linear shear-flow
which includes bubbles.  We predict that the generation of turbulence is
considerably suppressed when bubbles occupy a significant portion of the
flow in a channel, as in this study. At present, we believe that our
computations based on the assumption of laminar flows are reasonable and not
unrealistic.\\

%%10
\par\noindent
\item
\textsf
{Page 9, Fig. 5 and related discussion: It is quite strange that the bubble
does not reach a final steady shape, but seems to expand, contract and then
expand again. Here, it is even more important to verify that the simulations
have converged (via mesh and time refinement) and to extend them much further
in time to determine if this periodic motion will prevail or finally lead to a
steady state.}
\vspace{3 mm}

Response: \\
{\color{red}  
We have made a new figure (Figure 8: shear stress profile) for
the cases of 
(i) $Re$ = 50 and $Ca$ = 0.3, and
(ii) $Re$ = 93 and $Ca$ = 0.3.
Our new Figure 8 is a part of our new 
Section~I\hspace{-1.2pt}V.C.  The purpose of this new material
is to gain further insight on the unsteady dynamics of bubble
elongation.  What we discovered was that for the Reynolds
number sufficiently below the critical value, a relatively quick
unsteady elongation period gives way to a steady state (with no
break up).  On
the otherhand for Reynolds number {\em close} to the critical Reynolds
number, there is a prolonged, unsteady, elongation period, in which
periodic motion is observed and the deformation parameter $D$ is close 
to one.  It is agreed, that the ``vacillating''
behavior cannot last forever, ultimately (perhaps stochastically!),
the bubble will either settle down or break.  We assert that regardless
of the outcome, this vacillating behaviour will always occur
in close proximity to the critical Reynolds' number.  In otherwords,
regardless of the outcome, we claim that one is assured of being within 
3 percent of the critical Reynolds number (see Figure 3).  In fact,
we hypothesize that there will {\em always} be ``vacillating'' behavior
if one is sufficiently close to the critical Reynolds number.  i.e.
given an almost infinite supply of computational resources, as one hones
in closer and closer to the critical Reynolds number, a ``tug of war''
will be observed between the surface tension force trying to pull the
bubble together versus the wall driven shear stress trying to pull the
bubble apart.  This discussion
is added to the paper.
}
%%11
\par\noindent
\item
\textsf
{ Page 13: Fig. 11, which distinguishes the two areas of bubble breakup against
just bubble deformation, presents only 4 points and a curve connecting them,
although this is an important physical result. Several more points must be
included in it to reach a definitive conclusion.}
\vspace{3 mm}

Response: \\
We appreciate the reviewer's opinion, but it is important to keep in mind that
at the start of our study there were no guiding initial results that indicated
even potential values for critical Reynolds numbers for a bubble. The critical
Reynolds numbers found and presented were obtained after extensive and
expensive computations.  After lengthy computations, we were able to determine
the bubble $Re_{\rm c}$ number for 4 $Ca$ numbers.  We believe
that a bubble breakup critical curve drawn based on 4 points will be 
useful for low $Ca$ numbers.  Following the suggestion from the 
other reviewer (reviewer 1), 
we have added
a drop breakup critical curve ($\lambda$ = $\eta$ =1.0) 
to Fig. 12 (formerly Fig. 11).  
The difference in $Re_{\rm c}$ number between the bubble and 
the drop is clearer in
the new Fig. 12.  
In future work, we will determine the $Re_{\rm c}$ for a 
variety of high $Ca$ numbers 
in order to construct a detailed critical curve 
for a wider range of $Ca$ numbers.\\

%%12
\par\noindent
\item
\textsf
{\color{red}  End of conclusions section: .........
These phrases state the obvious, the densities and viscosities are different between drops and bubbles and, therefore, they cause different dynamics. 
Instead, it is imperative that the authors give the different physics that the property differences generate. 
To this end, they should examine thoroughly and in every detail the velocity and stress fields in each arrangement. 
This may require additional figures and text, but it must be done. Without this examination and its conclusions, this study is more appropriate for a 
Journal on numerical methods in fluid flows than a physical review Journal. This is the second most important issue with this presentation}
\vspace{3 mm}

### 2024-04-01 Bubble-deformation.tex
reply to reviewer 2 complete

{ \color{red}
The grid size and adaptive meshing strategy that we adopt is chosen
in order to answer the research question as to the conditions which determine
whether a bubble in shear flow
will break-up or not.  In such a case, we must accurately capture the 
balance of forces with respect to the (non-local) force exerted from the 
wall driven 
flow acting against the interfacial
surface tension force.  By nature of the research question, 
the accuracy of the transition
point for break-up depends on the 
largest Taylor Deformation parameter $D$
that is supported by the grid (see e.g. Figures \ref{fig:ShearStress} and
\ref{fig:BubBrkCa0p8Re43}).  
As we report here, we have found that as
long as the grid size is fine enough to support a Taylor Deformation parameter
$D<0.95$, then the transition region 
(see Figures \ref{fig:DeEvolution} and \ref{fig:CaRecFit}) will be captured
with a tolerance of three percent.  The simulation time becomes impractical
if we were to try to further improve the tolerance.  A smaller tolerance
would necessitate a longer supported Deformation
parameter $D$ which would in turn necessitate a higher aspect ratio
computational domain, increased droplet surface area at break-up,
increased number of time steps, and higher resolution for representing the
drop/bubble at its thinnest point.

@@

%% Ohta
{\color{blue} 
\subsection{Shear stress acting on the bubble}
%  -----------------------------------------------------------------------------
%
%\begin{comment} 
\begin{figure}%[h!]
  \centering
  \includegraphics[width=\textwidth]{ShearStress}
  \caption{Normalized shear stress $\tau^{\ast}$ around a bubble for 
	two Reynolds 
        numbers under the condition of $Ca = 0.3$. The left image shows the 
        shear stress profile at $Re=50$ and the right image at $Re=93$.
   }
  \label{fig:ShearStress}
\end{figure}
%\end{comment} 
%
In the previous section, the appearance of bubble deformation and breakup was
discussed. It is expected that a large deformation and breakup of the bubble
are closely related to the state of shear stress acting on the bubble.
Figure~\ref{fig:ShearStress} shows the shear stress profile around a bubble,
for two Reynolds numbers under the condition of $Ca = 0.3$.  The shear stress
profile on the left corresponds to the case of $Re = 50$ and the right side
shows the shear stress profile for the case of $Re = 93$.  The normalized shear
stress $\tau^{\ast} = \tau / \tau_{0}$ is defined as a ratio of the local shear
stress $\tau = \mu_{\rm m}$$\sqrt{2\bmD:\bmD}$ and the apparent shear stress
$\tau_{0} = \mu_{\rm m}$$\mathit{\Gamma}$.  In this study, for a given $Ca$
condition, the same value of $\tau_{0}$ is used regardless of $Re$. For the
case of $Re = 50$, the bubble reaches a deformed stable state.  As observed in
previous sections, when the value of $Re$ is slightly below the critical $Re$
condition, the bubble does not settle into a deformed stable state but instead
alternates in an elongation and contraction process.  In comparison, the right
image in Fig.~\ref{fig:ShearStress} shows a higher shear stress profile near
the bubble endpoints for the case of $Re = 93$ as it undergoes an elongation
state in the process toward breakup.  The value of the maximum shear stress for
the case of $Re = 50$ is $\tau ^{\ast} \approx 6$ and the maximum shear stress
for the case of $Re = 93$ at the moment shown in Fig.~\ref{fig:ShearStress} has
the value of $\tau ^{\ast} \approx 8$.  The shear stress profile of in
Fig.~\ref{fig:ShearStress} (color contour) is drawn in the range from $\tau
^{\ast} = 1$ to $\tau ^{\ast} = 8$, but, for emphasis, shear stress regions
with $\tau ^{\ast} \geqq 6$ are illustrated in red.  As can be seen in the
figure, the strongest shear stresses are concentrated on the ends of the bubble
for both $Re$ conditions.  This indicates that the strong shear stresses acting
on the ends of the bubble are responsible for much of the bubble stretching.
It is important to note that the magnitude of the shear stress acting on the
ends of the bubble for the case of $Re = 93$ is much larger than that for the
case of $Re = 50$.  On the other hand, the shear stress inside the bubble was
very small relative to that of the matrix fluid due to very small density and
viscosity of the bubble.  Since the force of strong shear stresses acting on
the ends of the bubble is difficult to transfer across the interface, as a
consequence, a sufficiently large $Re$ condition is required for large bubble
deformations.  }

>>>>

{ \color{red}
The grid size and adaptive meshing strategy that we adopt is chosen
in order to answer the research question as to the conditions which determine
whether a bubble in shear flow
will break-up or not.  In such a case, we must accurately capture the 
balance of forces with respect to the (non-local) force exerted from the 
wall driven 
flow acting against the interfacial
surface tension force.  
The accuracy of the ``Critical Reynolds Number''
depends on the 
largest Taylor Deformation parameter $D$
that is supported by the grid 
(see e.g. Figures \ref{fig:ShearStress} and
\ref{fig:BubBrkCa0p8Re43}).  
As we report here, we have found that as
long as the grid size is fine enough to support a Taylor Deformation parameter
$D<0.95$, then the transition region 
(i.e. ``Critical Reynolds number'')
(see Figures \ref{fig:DeEvolution} and \ref{fig:CaRecFit}) will be captured
with a tolerance of three percent.  The simulation time becomes impractical
if we were to try to further improve the ``critical Reynolds number''
accuracy.  
A smaller tolerance
would necessitate a larger supported Deformation
parameter $D$ which would in turn necessitate a higher aspect ratio
computational domain, increased droplet surface area at break-up,
increased number of time steps, and higher resolution for representing the
drop/bubble at its thinnest point.

@@

%% Ohta
{\color{blue} 
\subsection{Shear stress acting on the bubble}
%  -----------------------------------------------------------------------------
%
%\begin{comment} 
\begin{figure}%[h!]
  \centering
  \includegraphics[width=\textwidth]{ShearStress}
  \caption{Normalized shear stress $\tau^{\ast}$ around a bubble for 
	two Reynolds 
        numbers under the condition of $Ca = 0.3$. The left image shows the 
        shear stress profile at $Re=50$ and the right image at $Re=93$.
   }
  \label{fig:ShearStress}
\end{figure}
%\end{comment} 
%
In the previous section, the appearance of bubble deformation and breakup was
discussed. It is expected that a large deformation and breakup of the bubble
are closely related to the state of shear stress acting on the bubble.
Figure~\ref{fig:ShearStress} shows the shear stress profile around a bubble
for two Reynolds numbers under the condition of $Ca = 0.3$: Reynolds
number equal to 50 and 93.  
The shear stress
profile on the left corresponds to the case of $Re = 50$ and the right side
shows the shear stress profile for the case of $Re = 93$.  
The normalized shear
stress, $\tau^{\ast} = \tau / \tau_{0}$, is defined as a 
ratio of the local shear
stress $\tau = \mu_{\rm m}$$\sqrt{2\bmD:\bmD}$ 
and the apparent shear stress
$\tau_{0} = \mu_{\rm m}$$\mathit{\Gamma}$.  
In this study, for a given $Ca$ condition, the same 
value of $\tau_{0}$ is used regardless of $Re$. 
\par\noindent
For the
case of $Re = 50$, the bubble reaches a deformed stable state.  
\par\noindent
As observed in
previous sections, when the value of $Re$ is slightly below the 
critical $Re$ condition, the bubble does not settle into a deformed 
stable state but instead alternates in an elongation and contraction process.  
In comparison to the $Re=50$ case on the left, the right
image in Fig.~\ref{fig:ShearStress} ($Re=93$)
shows a higher shear stress profile near
the bubble endpoints as it undergoes an elongation
state in the process toward breakup.  
The value of the maximum shear stress for
the case of $Re = 50$ is $\tau ^{\ast} \approx 6$ 
and the maximum shear stress
for the case of $Re = 93$ at the moment 
shown in Fig.~\ref{fig:ShearStress} has
the value of $\tau ^{\ast} \approx 8$.  
The shear stress profile in
Fig.~\ref{fig:ShearStress} (color contour) is drawn in the range from 
$\tau^{\ast} = 1$ to $\tau ^{\ast} = 8$, but, for emphasis, 
shear stress regions
with $\tau ^{\ast} \geqq 6$ are illustrated in red.  
As can be seen in the
figure, the strongest shear stresses are concentrated on the 
ends of the bubble for both $Re$ conditions.  
This indicates that the strong shear stresses acting
on the ends of the bubble are responsible for much of the bubble stretching.
It is important to note that the magnitude of the shear stress acting on the
ends of the bubble for the case of $Re = 93$ is much larger than that for the
case of $Re = 50$.

We also observe that the shear stress {\em inside} the bubble was
very small relative to that of the matrix fluid due to very small density and
viscosity of the bubble.  Since the force of strong shear stresses acting on
the ends of the bubble is difficult to transfer across the interface, as a
consequence, a sufficiently large $Re$ condition is required for large bubble
deformations.  
\par\noindent
In summary, what we discover is that for the Reynolds
number sufficiently below the critical value, a relatively quick
unsteady elongation period gives way to a steady state (with no
break up).  On
the otherhand for Reynolds number {\em close} to the critical Reynolds
number, there is a prolonged, unsteady, elongation period, in which
periodic motion is observed and the deformation parameter $D$ is close 
to one.  The ``vacillating''
behavior cannot last forever, ultimately (perhaps stochastically!),
the bubble will either settle down or break.  We assert that regardless
of the outcome, this vacillating behaviour will always occur
in close proximity to the critical Reynolds' number.  In otherwords,
regardless of the outcome, we claim, using the grid resolution of
$R/16$, that one is assured of being within 
3 percent of the critical Reynolds number (see Figure \ref{fig:DeEvolution}).  
In fact,
we hypothesize that there will {\em always} be ``vacillating'' behavior
if one is sufficiently close to the critical Reynolds number.  i.e.
given an almost infinite supply of computational resources, as one hones
in closer and closer to the critical Reynolds number, a ``tug of war''
will be observed between the surface tension force trying to pull the
bubble together versus the wall driven shear stress trying to pull the
bubble apart.

}
